Search for Spatially Extended Fermi-LAT Sources Using Two Years of Flight Data

### 2011-07-24 extended_catalog/paper/extended_catalog.tex
adding stefan's revisions

% some units
\newcommand{\mev}{\text{MeV}\xspace}
\newcommand{\gev}{\text{GeV}\xspace}
\newcommand{\tev}{\text{TeV}\xspace}
\newcommand{\sr}{\text{sr}\xspace}
\newcommand{\s}{\text{s}\xspace}
\newcommand{\ph}{\text{ph}\xspace}
\newcommand{\cm}{\text{cm}\xspace}
\renewcommand{\sec}{\text{s}\xspace}
\newcommand{\tsext}{{\ensuremath{\text{TS}_\text{ext}}}\xspace}
\newcommand{\tsinc}{\ensuremath{\text{TS}_\text{inc}}\xspace}

@@

\begin{abstract}
We present a new method for analyzing spatially extended sources with
the Large Area Telescope (LAT), the primary science instrument
on the {\em Fermi Gamma-ray Space Telescope (Fermi)}.  We provide a
series of Monte Carlo studies to validate this tool
and calculate the LAT's detection threshold to spatially extended
sources.  We then apply this tool to test all sources in the two
year source catalog for extension. We present on the detection of
ten spatially extended sources in addition to the twelve spatially
extended sources reported in the second Fermi LAT Catalog.
\end{abstract}

@@

The Large Area Telescope (LAT) is a pair conversion telescope on the
Fermi Gamma-Ray Space Telescope. It has been surveying the $\gamma$-ray
sky since June 2008.  Fermi is well suited as a survey
instrument because of its broad energy coverage (20 \mev
to $>300$ \gev), wide field of view ($\sim 2.4 \sr$), and large effective
area ($\sim 8000 \cm^2$ at $>1 \gev$).

@@

Morphological studies of sources in the \gev energy range
using the LAT are challenging because of the significantly energy
dependent PSF and because of systematic errors in the galactic diffuse
emission.  The LAT's PSF is determined at lower energies by multiple
scattering in the tracker and the 68\% containment radius of the PSF is
approximately 5.1\deg at 100 \mev  when averaged over instrument acceptance
and including photons which convert in either the thick or thin layers of
the tracker. The PSF approaches a limit given by the granularity of the
tracker channels and is 0.14\deg at 10 \gev (\cite{on_orbit_calibration}).
Therefore, the higher energy photons are significantly more important for
resolving a source but are typically limited by statistics. 
A detailed analysis procedure must be used to maximize our
sensitivity over all energies.

@@

This analysis is not feasible using the official LAT likelihood
analysis tool \gtlike\footnote{\gtlike is distributed publicly by the
Fermi Science Support Center (\cite{fssc})} because \gtlike can only 
fit the spectral parameters of the model. The extension fitting would
have to be done as an external loop. Typically, the
run time for a binned source analysis using \gtlike is several hours
so this loop would be especially time consuming. What has typically
been done to study an extended sources is to fix the position of the
source and develop a profile of the likelihood as a
function of extension.  This approach is not optimal because the
position and extension of the source should be simultaneously maximized
to find the best fit parameters and to correctly compute the statistical
significance of the detection.  Furthermore since this approach is
computationally intensive, no large scale Monte Carlo effort can
be run to validate it.

@@

What differentiates \pointlike most significantly from \gtlike is its
emphasis on speed by introducing approximations into the calculation
of the likelihood function while trying to minimized the introduced
numerical errors.  Like the binned version of \gtlike, \pointlike bins
the sky in both position and energy.  Unlike \gtlike, \pointlike relies on
the {\em healpix} representation of spatial bins (\cite{healpix}). Since
the PSF changes significantly with energy, the spatial size of each bin
is calculated to be smaller than than the PSF. By changing the spatial
bin size with energy, \pointlike ensures that little information is lost
at each energy while minimizing the required number of bins.

\pointlike also bins $\gamma$-rays depending upon whether they convert
in the thin or thick layers of the tracker. Since the PSF is smaller for
front entering photons, by binning each conversion type independently,
\pointlike improves the overall sensitivity.  Furthermore, \pointlike
uses a sparse matrix representation of the spatial bins where only the
bins with counts in them are used. To evaluate the likelihood, \pointlike
then independently integrates the model predicted counts in the region.

@@

Finally, \pointlike is written mostly in the \python programming
language and uses {\em numpy} and {\em scipy} to vectorize most of the
calculations. This provides a reasonable trade off between computational
efficiency and ease of development. By leaving the interface as a series
of \python modules, advanced functions including extension fitting
(section~\ref{extension_fitting}) and dual localization
(section~\ref{dual_localization_method}) could be quickly
built on top of the core functionality.

@@

We modified \pointlike to allow modeling the spatial extension 
of a source. It is assumed that the spatial and spectral information
completely decouple and that the spatial shape can be represented by a
simple geometric function.  The code consistently convolves the extended
source shape with the PSF (as a function of energy) when computing the
model predicted counts.

\pointlike uses the \minuit fitting library to maximize the likelihood
by simultaneously fitting the position and extension of an extended
source (\cite{minuit_documentation}).  As will be described in
section~\ref{monte_carlo_validation}, this simultaneous fitting
is necessary to correctly calculating the statistical significance
of the detection of extension.
For each position and extension,
the spectral parameters of the sky model are refit. To avoid projection
effects, what is directly fit is not the longitude and latitude of
the source but instead the displacement in a rotated reference frame.

The significance of the extension of a source can be calculated from
\tsext, which is defined as twice the increase in log likelihood between
the extended source model at its best fit values and the point source
model at its best fit values:
\begin{equation}
  \tsext=2\log(LL_\text{ext}/LL_\text{ps}).
\end{equation}
\pointlike calculates \tsext by fitting a source first with a spatially
extended model and then as a point source.  The interpretation
of \tsext in terms of a statistical significance is discussed in
section~\ref{monte_carlo_validation}.

@@

We found that the {\em HESSE} errors on the extension of a source
were sometimes badly estimated.  On the other
hand, the {\em Simplex} algorithm was prohibitively slow and would often
fail to convergence. Our approach to calculating extension errors
is to fix the position of the source and varying just the extension
using the {\em Simplex} algorithm.  This approach
makes the approximation that the covariance between each spatial parameter
is small but not that the likelihood is Gaussian. This method 
is shown schematically in figure~\ref{extension_profile} which shows
the change in log likelihood when varying the extension of
the SNR IC443.
The localization error separately calculated by fixing the extension to the
best fit value and fitting an ellipse to the likelihood as a function
of the spatial parameters. This
elliptical localization error correctly accounts for the covariance
in the position of the source and this same me method is used to
calculate the localization error of all point sources in the first LAT catalog
(\cite{first_cat}).

@@

There is a degeneracy between a spatially extended source and multiple
point sources.  To assess the possibility of source confusion,
we developed a function in \pointlike to simultaneously fit 
the position of two point sources.
The code uses \minuit to fit the position of the two point
sources.  To avoid projection effects, the fit is done in a rotated coordinate
system.  To improve the robustness of the fit, Instead of directly fitting
the rotated coordinates $x_1$ and $x_2$ we instead
fit the sum and difference of the rotated
coordinates $S=(x_1+x_2)/2$ and $D=(x_2-x_1)/2$.  The log likelihood
is then maximized by fitting the sum and difference of the 
sources.  Because of \pointlike's computational efficiency, this algorithm
runs in a reasonable amount of time.

We define \tsinc as twice the increase in log likelihood
fitting the region as two point sources compared to fitting the region
as one point source:
\begin{equation}
  \tsinc=2\times(LL_\text{2pts}-LL_\text{ps}).
\end{equation}
\tsinc can not be directly compared to \tsext to see which model is 
more significant
because the models are not nested (\cite{statistics_with_care}). Even
though the comparison of \tsext with \tsinc is not a calibrated test,
we find the extreme cases $\tsinc \ll \tsext$ or $\tsinc\gg\tsext$
to be suggestive and we only consider a source to be extended if
$\tsext>\tsinc$.

@@

% 6 SNRs
The largest population of extended source are SNRs. The six extended SNRs
included in the catalog are IC443, W28, W30, W44, W51C, and the Cygnus
Loop. These sources were first reported in (\cite{ic443,w28,w44,w51c}).
Detailed papers reporting on W30 and the Cygnus Loop are still in
preparation.  Using photons with energies greater than 1 \gev, our
analysis significantly detected the extension of the SNRs IC443, W28,
W30, W44, and W51C.

Our analysis failed to correctly fit the extension of the Cygnus Loop.
Publications describing the detailed analysis for the Cygnus Loop are
still in preparation.  This is most likely because Cygnus Loop is rather
large and its shell type morphology is not well represented by a uniform
disk. The source is also rather soft so the published results used a
larger energy range then our search.

@@

% 3 PWN
Three pulsar wind nebula Vela X, MSH 15-52, and HESS J1825-137 
were included in the second catalog (\cite{velax,msh1552,fermi_hess_j1825}).

Our analysis failed to detect the extension of Vela X. Vela X can only be
successfully analyzed after applying a phase cut to removed the heavily
dominant pulsar emission. It is beyond the scope of the paper to perform
such specialized analysis.

Using photons with energy above 10 \gev, we significantly detected
the extension of HESS J1825-137.
Our analysis of MSH 15-52 using photons with energy above 10 \gev found an extension
consistent with the published result, but with $\tsext=11.2$. This is
most likely because the published analysis used a larger energy range,
considering all events above 6 \gev. Looking at all photons with
energy above 1 \gev 
is problematic because of background source confusion whereas using photons
with energy above 10 \gev does not have the statistics for a significant detection.

\todo[inline]{note about how MSH 15-52 and Vela X were hard
to analysze without phase cuts.}

@@

% Notes about extended Catalog sources
% Catalog used for analysis is P72Y_uw23.fits
% Compare to final catalog gll_psc_v04.fit
%          1FGL Name - Preliminary 2FGL - 2FGL Name
% 1FGL J1628.6-2419c - P72Y2516         - 2FGL J1627.0-2425c
% 1FGL J1554.0-5345c - P72Y2405         - 2FGL J1554.4-5317c
% 1FGL J0823.3-4248  - P72Y1212         - 2FGL J0823.0-4246
% 1FGL J1613.6-5100c - P72Y2472         - 2FGL J1615.0-5051 
% 1FGL J1614.7-5138c - P72Y2473         - 2FGL J1615.2-5138
% 1FGL J1632.9-4802c - P72Y2540         - 2FGL J1632.4-4753c
% 1FGL J1711.7-3944c - P72Y2674         - 2FGL J1712.4-3941
% 1FGL J2020.0+4049  - P72Y3281         - 2FGL J2021.5+4026
% 1FGL J1837.5-0659c - P72Y2974         - 2FGL J1837.3-0700c
% N/A                - P72Y1287         - 2FGL J0851.7-4635

@@

% (no 1FGL) - P72Y1287         - 2FGL J0851.7-4635
% Delete:
% * P72Y1291 - 2FGL J0853.5-4711
% * P72Y1296 - 2FGL J0855.4-4625
% * P72Y1274 - 2FGL J0848.5-4535
% * P72Y1259 - 2FGL J0842.9-4721
% * P72Y1300 - 2FGL J0858.0-4815
% Modify:
% * P72Y1293 - 2FGL J0854.7-4501 - SkyDir(266.232,0.517,SkyDir.GALACTIC))
%              (initial Position - SkyDir(265.5705,-0.0099)
% * P72Y1310 - 2FGL J0901.7-4655 - SkyDir(267.311,-0.190,SkyDir.GALACTIC)
%              (initial Position - SkyDir(267.79,-0.3339)

@@

To get a good fit of the source, we had to remove three nearby catalog
sources 2FGL J0848.5-4535, 2FGL J0855.4-4625, and 2FGL J0853.5-4711
which are part of the extended emission.  In addition, we relocalized
the position of two nearby catalog sources 2FGL J0854.7-4501 and 2FGL
J0901.7-4655 to better fit their position at high energies in the presence
of an extended source.  This localization was done iteratively alternately 
fitting the position of these two point sources and the extension of
2FGL J0851.7-4635 until the fit globally converged.  2FGL J0854.7-4501
moved to $(l,b)=(266.232\deg,0.517\deg)$ and 2FGL J0901.7-4655 moved to
$(l,b)=(267.311\deg,-0.190\deg)$. In addition, we removed the further away
catalog source 2FGL J0858.0-4815 because it was not significant above 10
\gev.  After modifying the region, we found that \tsinc=16.2 which is less
than \tsext so we conclude that the source is extended 
These modifications are shown in
figure~\ref{Vela_Jr}.

@@

This source is less than 1\deg away from 2FGL J1615.0-5051 which is
also spatially extended (see section~\ref{section_2FGL_J1615.2-5138}).
To get a good fit of both sources, we had to model both sources as being
spatially extended and iteratively fit the position and extension of each
source until obtaining a global best fit.  Before doing this, We removed
from our model the three nearby catalog sources 2FGL J1620.6-5111c, 2FGL
J1619.7-5040c, and 2FGL J1622.8-5006 2FGL J1615.0-5051 because they were
not significant above 10 \gev.  These modifications are further described
in the caption to figure~\ref{1FGL_J1613.6-5100c}.  After modifying
the region, we found that \tsinc=12.2 which is less than \tsext so we
conclude that the source is extended.

This source is spatially coincident with the extended \tev source HESS
J1616-508 (\cite{HESS_plane_survey}).  In figure~\ref{1FGL_J1613.6-5100c},
contours of HESS J1616-508 are overlaid on 2FGL J1615.0-5051.  The
H.E.S.S. experiment measured an extension $\sigma$ of $0.23\pm0.02$ and
$0.15\pm0.02$ along the semi-major and semi-minor axes respectively when
fitting the source with an elliptical Gaussian spatial model.  This size
corresponds to a 68\% source containment radius of $\rsixeight\sim0.22\deg$
to $0.34\deg$. This size matches very well with the LAT size
$\rsixeight=0.27\deg\pm0.03$ (see section~\ref{compare_source_size}).

@@

This source is spatially coincident with the extended \tev source
HESS J1614-518 (\cite{HESS_plane_survey}). In figure~\ref{1FGL_J1613.6-5100c},
contours of HESS J1614-518 are overlaid on 2FGL J1615.2-5138.  The H.E.S.S. experiment measured a Gaussian extension of
$\sigma=0.156\deg\pm0.008\deg$. This size corresponds to a 68\% source
containment radius of $\rsixeight\sim0.33\pm0.01\deg$.  This size matches
very well with the LAT size of $\rsixeight\sim0.34\pm0.02$.

@@

To test for source confusion, we compared the likelihood fitting this
region as one extended source to the likelihood fitting this region as
two point sources. We found $\tsinc=40.9\ll\tsext$ which is less than \tsext
so we conclude that this is an extended source.

@@

This is RX J1713.7-2946. This analysis not finished.

Recent LAT publication: \cite{rx_j1713_lat}.

Contours taken from HESS Paper: \cite{rx_j1713_hess}

The Figure is \ref{2FGL_J1712.4-3941}

@@

@@

\clearpage
\begin{figure}
  \begin{center}
    \includegraphics{ic443_plots/profile_ic443.pdf}
    % taken from /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/extension_profile_ic443/v1/profile_ic443.pdf
    \caption{
    A plot of the overall change in \ts when varying the
    extension of the extended SNR IC443.  As is discussed in section
    \ref{extension_error}, the error on the extension of a source is
    calculated by varying the extension away from its best fit value until
    the log likelihood has changed by a particular amount.  These profiles are
    generated for all extended source candidates. They are useful
    to make sure the fitter find a global minimum and that
    the likelihood function is not unphysically double peaked.
    }
    \label{extension_profile}
  \end{center}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \includegraphics[type=pdf,ext=.pdf,read=.pdf]{source_plots/source_1FGL_J0823.3-4248}
  \end{center}
  % this plot came from 
  % /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/source_plots/1FGL_J0823.3-4248/v2/source_1FGL_J0823.3-4248.pdf
  \caption{A galactic and isotropic diffuse subtracted 1 \gev to 100
  \gev counts map of the region around 2FGL J0823.0-4246 smoothed by a
  0.1\deg Gaussian kernel.  The red star is the catalog position of this source.  
  The red cross and circle are the best fit position
  and extension of this source assuming a uniform disk spatial model.
  The two green stars are the positions of the sources 2FGL J0823.4-4305
  and 2FGL J0821.0-4254 which were removed because they are part of the
  extended source.  This source is spatially coincident with the Puppis
  A SNR. The green contours are an X-ray image of Puppis A observed by
  ROSAT (\cite{rosat_puppis_a}). There is good spatial overlap between
  the LAT and X-ray emission.  Even though the Vela pulsar is over 3
  degrees away from the source, because it is so bright it can be seen
  on the top left.  More information about this source can be found in 
  section~\ref{section_2FGL_J0823.0-4246}.
  }\label{1FGL_J0823.3-4248}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[type=pdf,ext=.pdf,read=.pdf]{source_plots/source_1FGL_J1628.6-2419c}
  \end{center}
  % this plot came from 
  % /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/source_plots/1FGL_J1628.6-2419c/v2/run.py
  \caption{
  A galactic and isotropic diffuse subtracted 1 \gev to 100 \gev counts
  map of the region around 2FGL J1627.0-2425 smoothed by a 0.1\deg
  Gaussian kernel.  The red star is the catalog's position of
  this source.  The red cross and circle are
  the best position and extension of this source assuming a
  uniform disk spatial model.  The black cross is the background source
  2FGL J1625.7-2526.  The green contours are a 100 micrometer infrared
  image of the region as seen by IRAS (\cite{iras_rho_ophiuci}). This
  source is in the Ophiuchus region and most likely represents 
  unmodeled diffuse emission.  More information about this source can be found
  in section~\ref{section_2FGL_J1627.0-2425c}.
  }\label{1FGL_J1628.6-2419c}
\end{figure}

>>>>

% some units
\newcommand{\mev}{\text{MeV}\xspace}
\newcommand{\gev}{\text{GeV}\xspace}
\newcommand{\tev}{\text{TeV}\xspace}
\newcommand{\sr}{\text{sr}\xspace}
\newcommand{\s}{\text{s}\xspace}
\newcommand{\ph}{\text{ph}\xspace}
\newcommand{\cm}{\text{cm}\xspace}
\renewcommand{\sec}{\text{s}\xspace}
\newcommand{\tsext}{{\ensuremath{\text{TS}_\text{ext}}}\xspace}
\newcommand{\tsinc}{\ensuremath{\text{TS}_\text{inc}}\xspace}
\newcommand{\loglikelihood}{\ensuremath{LL}\xspace}

@@

\begin{abstract}
We present a new method for analyzing spatially extended sources with
the Large Area Telescope (LAT), the primary science instrument on the
{\em Fermi Gamma-ray Space Telescope (Fermi)}. Measuring the extension
of a LAT source is important for uniquely identifying it.  We provide
a series of Monte Carlo studies to validate this tool and calculate
the LAT's detection threshold to spatially extended sources.  We then
apply this tool to test all sources in the two year source catalog for
extension. We present on the detection of ten spatially extended sources
in addition to the twelve spatially extended sources reported in the
second Fermi LAT Catalog.
\end{abstract}

@@

The Large Area Telescope (LAT) is a pair conversion telescope on the
Fermi Gamma-Ray Space Telescope (Fermi). It has been surveying the $\gamma$-ray
sky since June 2008.  Fermi is well suited as a survey
instrument because of its broad energy coverage (20 \mev
to $>300$ \gev), wide field of view ($\sim 2.4 \sr$), and large effective
area ($\sim 8000 \cm^2$ at $>1 \gev$).

@@

Morphological studies of sources in the \gev energy range
using the LAT are challenging because of the significantly 
energy-dependent PSF and because of systematic errors in the galactic diffuse
emission.  The LAT's PSF is determined at lower energies by multiple
scattering in the tracker and the 68\% containment radius of the PSF is
approximately 5.1\deg at 100 \mev  when averaged over instrument acceptance
and including photons which convert in either the thick or thin layers of
the tracker. The PSF approaches a limit given by the granularity of the
tracker channels and is 0.14\deg at 10 \gev (\cite{on_orbit_calibration}).
Therefore, the higher energy photons are significantly more important for
resolving a source but are typically limited by statistics. 
A detailed analysis procedure must be used to maximize our
sensitivity over all energies.

@@

This analysis is not feasible using the official LAT likelihood
analysis tool \gtlike\footnote{\gtlike is distributed publicly by the
Fermi Science Support Center (\cite{fssc})} because \gtlike can only 
fit the spectral parameters of the model. The extension fitting would
have to be done as an external loop which would typically be
too time consuming. What has typically
been done to study an extended sources is to fix the position of the
source and develop a profile of the likelihood as a
function of extension.  This approach is not optimal because the
position and extension of the source should be simultaneously maximized
to find the best fit parameters and to correctly compute the statistical
significance of the detection.  Furthermore since this approach is
computationally intensive, no large scale Monte Carlo effort can
be run to validate it.

@@

What differentiates \pointlike most significantly from \gtlike is its
emphasis on speed by introducing approximations into the calculation
of the likelihood function while trying to minimized the introduced
numerical errors.  Like the binned version of \gtlike, \pointlike bins
the sky in both position and energy.  Unlike \gtlike, \pointlike relies on
the {\em healpix} representation of spatial bins (\cite{healpix}). Since
the PSF changes significantly with energy, the spatial size of each bin
is chosen to be smaller than the PSF. By changing the spatial
bin size with energy, \pointlike ensures that little information is lost
at each energy while minimizing the required number of bins.

\pointlike also bins $\gamma$-rays depending upon whether they convert
in the thin or thick layers of the tracker. Since the PSF is smaller for
photons which convert in the thin layers of the tracker,
photons are also binned depending upon where they 
convert in the tracker.
Furthermore, \pointlike
uses a sparse matrix representation of the spatial bins where only the
bins with counts in them are used. To evaluate the likelihood, \pointlike
then independently integrates the model predicted counts in the region.

@@

@@

In \pointlike, it is assumed that the spatial and spectral model
of an extended source decouple and that the spatial shape can be
represented by a two dimensional function.  To fit the extended source,
\pointlike convolves the extended source shape with the PSF (as a
function of energy) and uses the \minuit fitting library to maximize
the likelihood by simultaneously fitting the position and extension
of the source (\cite{minuit_documentation}).  As will be described in
section~\ref{monte_carlo_validation}, simultaneously fitting the position
and extension is necessary to correctly calculate the statistical
significance of the detection of extension.  For each position and
extension, the spectral parameters of the sky model are refit. To avoid
projection effects, what is directly fit is not the longitude and latitude
but instead the source's displacement in a rotated reference frame.

The significance of the extension of a source can be calculated from
\tsext, which is defined as twice the increase in log likelihood (\loglikelihood)
between
the extended source model at its best fit values and the point source
model at its best fit values:
\begin{equation}
  \tsext=2\log(\loglikelihood_\text{ext}/\loglikelihood_\text{ps}).
\end{equation}
\pointlike calculates \tsext by fitting a source first with a spatially
extended model and then as a point source.  The interpretation
of \tsext in terms of a statistical significance is discussed in
section~\ref{monte_carlo_validation}.

@@

We found that the {\em HESSE} errors on the extension of a source
were sometimes badly estimated.  On the other
hand, the {\em Simplex} algorithm was prohibitively slow and would often
fail to convergence. Our approach to calculating extension errors
is to fix the position of the source and varying just the extension
using the {\em Simplex} algorithm.  This approach
makes the approximation that the covariance between each spatial parameter
is small but not that the likelihood is Gaussian. This method 
is shown schematically in figure~\ref{extension_profile} which shows
the change in \loglikelihood when varying the extension of
the SNR IC443.
The localization error separately calculated by fixing the extension to the
best fit value and fitting an ellipse to the likelihood as a function
of the spatial parameters. This
elliptical localization error correctly accounts for the covariance
in the position of the source and this same me method is used to
calculate the localization error of all point sources in the first LAT catalog
(\cite{first_cat}).

@@

There is a degeneracy between a spatially extended source and multiple
point sources.  To assess the possibility of source confusion,
we developed a function in \pointlike to simultaneously fit 
the position of two point sources.
The code uses \minuit to fit the position of the two point
sources.  To avoid projection effects, the fit is done in a rotated coordinate
system.  To improve the robustness of the fit, Instead of directly fitting
the rotated coordinates $x_1$ and $x_2$ we instead
fit the sum and difference of the rotated
coordinates $S=(x_1+x_2)/2$ and $D=(x_2-x_1)/2$.  The \loglikelihood
is then maximized by fitting the sum and difference of the 
sources.  Because of \pointlike's computational efficiency, this algorithm
runs in a reasonable amount of time.

We define \tsinc as twice the increase in \loglikelihood
fitting the region as two point sources compared to fitting the region
as one point source:
\begin{equation}
  \tsinc=2\times(\loglikelihood_\text{2pts}-\loglikelihood_\text{ps}).
\end{equation}
\tsinc can not be directly compared to \tsext to see which model is 
more significant
because the models are not nested (\cite{statistics_with_care}). Even
though the comparison of \tsext with \tsinc is not a calibrated test,
we find the extreme cases $\tsinc \loglikelihood \tsext$ or $\tsinc\gg\tsext$
to be suggestive and we only consider a source to be extended if
$\tsext>\tsinc$.

@@

% 6 SNRs
The largest population of extended source are SNRs. The six
extended SNRs included in the second catalog are IC443, W28, W30,
W44, W51C, and the Cygnus Loop. These sources were first reported in
(\cite{ic443,w28,w44,w51c}).  Detailed papers studying W30 and the Cygnus
Loop are still in preparation.  Using photons with energies between 1
\gev and 100 \gev, our analysis significantly detected the extension
of all six SNRs. Table~\ref{known_extended_sources} shows the best fit
positions and extensions of these sources.

@@

% 3 PWN
Three pulsar wind nebula Vela X, MSH 15-52, and HESS J1825-137 were
included in the second catalog (\cite{velax,msh1552,fermi_hess_j1825}).
Using photons with energy above 10 \gev, we significantly detected the
extension of HESS J1825-137.  To improve our model of this source, we
removed the nearby catalog source 2FGL J1823.1-1338c which is part of the
extended emission of HESS J1825-137.  Using photons with energies above
10 \gev, we fit the extension of MSH 15-52 consistent with the published
result at an extension significance of \tsext=9.5.  Our analysis was
unable to resolve Vela X. An analysis of this source requires removing on
pulse events from the Vela pulsar and is beyond the scope of this paper.

\todo[inline]{note about how MSH 15-52 can't go lower in
energy because of nearby pulsar.}

@@

% Notes about extended Catalog sources
% Catalog used for analysis is P72Y_uw23.fits
% Compare to final catalog gll_psc_v04.fit
%          1FGL Name - Preliminary 2FGL - 2FGL Name
% 1FGL J1628.6-2419c - P72Y2516         - 2FGL J1627.0-2425c
% 1FGL J1554.0-5345c - P72Y2405         - 2FGL J1554.4-5317c
% 1FGL J0823.3-4248  - P72Y1212         - 2FGL J0823.0-4246
% 1FGL J1711.7-3944c - P72Y2674         - 2FGL J1712.4-3941

% 1FGL J1613.6-5100c - P72Y2472         - 2FGL J1615.0-5051 
% 1FGL J1614.7-5138c - P72Y2473         - 2FGL J1615.2-5138
% 1FGL J1632.9-4802c - P72Y2540         - 2FGL J1632.4-4753c
% 1FGL J2020.0+4049  - P72Y3281         - 2FGL J2021.5+4026
% 1FGL J1837.5-0659c - P72Y2974         - 2FGL J1837.3-0700c
% N/A                - P72Y1287         - 2FGL J0851.7-4635

@@

% (no 1FGL) - P72Y1287         - 2FGL J0851.7-4635
% Delete:
% * P72Y1291 - 2FGL J0853.5-4711
% * P72Y1296 - 2FGL J0855.4-4625
% * P72Y1274 - 2FGL J0848.5-4535
% * P72Y1259 - 2FGL J0842.9-4721
% * P72Y1300 - 2FGL J0858.0-4815
% Modify:
% * P72Y1293 - 2FGL J0854.7-4501 - SkyDir(266.235,0.493,SkyDir.GALACTIC))
%              (initial Position - SkyDir(265.5705,-0.0099)
% * P72Y1310 - 2FGL J0901.7-4655 - SkyDir(267.575,-0.191,SkyDir.GALACTIC)
%              (initial Position - SkyDir(267.79,-0.3339)

@@

To get a good fit of the source, we had to remove three nearby catalog
sources 2FGL J0848.5-4535, 2FGL J0855.4-4625, and 2FGL J0853.5-4711
which are part of the extended emission.  In addition, we relocalized
the position of two nearby catalog sources 2FGL J0854.7-4501 and 2FGL
J0901.7-4655 to better fit their position at high energies in the presence
of an extended source.  This localization was done iteratively alternately 
fitting the position of these two point sources and the extension of
2FGL J0851.7-4635 until the fit globally converged.  2FGL J0854.7-4501
moved to $(l,b)=(266.24\deg,0.49\deg)$ and 2FGL J0901.7-4655 moved to
$(l,b)=(267.58\deg,-0.19\deg)$. In addition, we removed the further away
catalog source 2FGL J0858.0-4815 because it was not significant above 10
\gev.  After modifying the region, we found that \tsinc=16.2 which is less
than \tsext so we conclude that the source is extended 
These modifications are shown in
figure~\ref{Vela_Jr}.

@@

This source is less than 1\deg away from 2FGL J1615.2-5138 which is
also spatially extended (see section~\ref{section_2FGL_J1615.2-5138}).
To get a good fit of both sources, we had to model both sources as
being spatially extended and iteratively fit the position and extension
of each source until obtaining a global best fit.  Before doing this,
We removed from our model the source 2FGL J1614.9-5212 because it is
part of 2FGL J1615.2-5138's extended emission. Furthermore, we removed
the nearby catalog sources 2FGL J1619.7-5040c and 2FGL J1620.6-5111c
because they were not significant above 10 \gev.  These modifications
are further described in the caption to figure~\ref{1FGL_J1613.6-5100c}.
After modifying the region, we found that \tsinc=12.2 which is less than
\tsext so we conclude that the source is extended.

This source is spatially coincident with the extended \tev source HESS
J1616-508 (\cite{HESS_plane_survey}).  In figure~\ref{1FGL_J1613.6-5100c},
contours of HESS J1616-508 are overlaid on 2FGL J1615.0-5051.
The H.E.S.S. experiment measured an extension $0.136\pm 0.008$
when fitting the source with an elliptical Gaussian spatial model.
This size corresponds to a 68\% source containment radius of
$\rsixeight\sim0.21\deg\pm0.01$. This size is comparable to the LAT size
$\rsixeight=0.27\deg\pm0.03$ (see section~\ref{compare_source_size}).

@@

This source is spatially coincident with the extended \tev source HESS
J1614-518 (\cite{HESS_plane_survey}). In figure~\ref{1FGL_J1613.6-5100c},
contours of HESS J1614-518 are overlaid on 2FGL J1615.2-5138.
The H.E.S.S. experiment measured a Gaussian extension of
$\sigma=0.23\deg\pm0.02\deg$ and $\sigma=0.15\pm0.02$ in the semi-major
and semi-minor axis. This size corresponds to a 68\% containment size
of $\rsixeight=0.35\pm0.03\deg$ and $0.23\pm0.03$.  This elliptical size
matches well the LAT size of $\rsixeight\sim0.34\pm0.02$.

@@

To test for source confusion, we compared the likelihood fitting this
region as one extended source to the likelihood fitting this region as
two point sources. We found $\tsinc=40.9\loglikelihood\tsext$ which is less than \tsext
so we conclude that this is an extended source.

@@

The two year catalog source 2FGL J1712.4-3941 was found to be
spatially extended using photons with energies between 10 \gev and
100 \gev. The source is coincident with the one year catalog source
1FGL J1711.7-3944c. We fit an extension of $0.56\deg\pm0\pm04$ with an
extension significance of $\tsext=41.2$. This source was fit to a position
of $(l,b)=(347.25\deg,-0.54\deg)$. Figure \ref{2FGL_J1712.4-3941} shows
a smoothed counts map of this source.

This source is spatially conincident with the SNR RX J1713.7-2946 and was
recently reported by (\cite{rx_j1713_lat}).  To analyze this source,
we used the same background model as the recent LAT publication.
2FGL J1715.4-4024c is spatially coincident with Source A and was
moved to the published location of $(ra,dec)=(258.84,-40.46)$. Source
B was added at $(ra,dec)=(258.71,-38.70)$ and Source C was added at
$(ra,dec)=(257.47,-39.75)$.  Figure \ref{2FGL_J1712.4-3941} shows
the position of these nearby sources on top of a contour of SNR RX
J1713.7-2946 seen in \tev by H.E.S.S (\cite{rx_j1713_hess}).

@@

\todo[inline]{Describe comparison of \gev and \tev extensions from figure~\ref{gev_vs_tev_histogram}}

@@

\clearpage
\begin{figure}
  \begin{center}
    \includegraphics{ic443_plots/profile_ic443.pdf}
    % taken from /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/extension_profile_ic443/v1/profile_ic443.pdf
    \caption{
    A plot of the overall change in \ts when varying the
    extension of the extended SNR IC443.  As is discussed in section
    \ref{extension_error}, the error on the extension of a source is
    calculated by varying the extension away from its best fit value until
    the \loglikelihood has changed by a particular amount.  These profiles are
    generated for all extended source candidates. They are useful
    to make sure the fitter find a global minimum and that
    the likelihood function is not unphysically double peaked.
    }
    \label{extension_profile}
  \end{center}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \includegraphics[type=pdf,ext=.pdf,read=.pdf]{source_plots/source_1FGL_J0823.3-4248}
  \end{center}
  % this plot came from 
  % /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/source_plots/1FGL_J0823.3-4248/v2/source_1FGL_J0823.3-4248.pdf
  \caption{A galactic and isotropic diffuse subtracted 1 \gev to 100
  \gev counts map of the region around 2FGL J0823.0-4246 smoothed by a
  0.1\deg Gaussian kernel.  The red star is the catalog position of this source.  
  The red cross and circle are the best fit position
  and extension of this source assuming a uniform disk spatial model.
  The two green stars are the positions of the sources 2FGL J0823.4-4305
  and 2FGL J0821.0-4254 which were removed because they are part of the
  extended source.  This source is spatially coincident with the Puppis
  A SNR. The light blue contours are an X-ray image of Puppis A observed by
  ROSAT (\cite{rosat_puppis_a}). There is good spatial overlap between
  the LAT and X-ray emission.  Even though the Vela pulsar is over 3
  degrees away from the source, because it is so bright it can be seen
  on the top left.  More information about this source can be found in 
  section~\ref{section_2FGL_J0823.0-4246}.
  }\label{1FGL_J0823.3-4248}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[type=pdf,ext=.pdf,read=.pdf]{source_plots/source_1FGL_J1628.6-2419c}
  \end{center}
  % this plot came from 
  % /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/source_plots/1FGL_J1628.6-2419c/v2/run.py
  \caption{
  A galactic and isotropic diffuse subtracted 1 \gev to 100 \gev counts
  map of the region around 2FGL J1627.0-2425 smoothed by a 0.1\deg
  Gaussian kernel.  The red star is the catalog's position of
  this source.  The red cross and circle are
  the best position and extension of this source assuming a
  uniform disk spatial model.  The black cross is the background source
  2FGL J1625.7-2526.  The light blue contours are a 100 micrometer infrared
  image of the region as seen by IRAS (\cite{iras_rho_ophiuci}). This
  source is in the Ophiuchus region and most likely represents 
  unmodeled diffuse emission.  More information about this source can be found
  in section~\ref{section_2FGL_J1627.0-2425c}.
  }\label{1FGL_J1628.6-2419c}
\end{figure}
